---
title: "Analysis of sHLA-G levels"
author: "Sushrut Vyawahare"
date: "7 August 2019"
output: html_document
---

### Loading Libraries
```{r}
library(dplyr)
library(lme4)
library(ggplot2)
library(ggpubr)
library(gamlss)

```

### Loading AGA Data
```{r}
AGA <- read.csv("C:/Users/DELL/Downloads/AGA1.csv")
head(AGA)
AGA$Group <- "AGA"

```

```{r}
names(AGA) <- c("Enrid","POG","SHLA.G.levels","Vist","Group")
head(AGA)

```

### EDA Process For AGA

#### Missing values
```{r}
sum(is.na(AGA))
AGA <- na.omit(AGA)

```

#### Removing outliers
```{r}
boxplot(AGA$SHLA.G.levels)
boxplot(AGA$SHLA.G.levels)$out
boxplot(AGA$SHLA.G.levels, plot=FALSE)$out  
outliers <- boxplot(AGA$SHLA.G.levels, plot=FALSE)$out  
AGA[which(AGA$SHLA.G.levels %in% outliers),]
AGA <- AGA[-which(AGA$SHLA.G.levels %in% outliers),]

```

#### Boxplot after removing outliers
```{r}
boxplot(AGA$SHLA.G.levels)

```

#### Log transformation as data is not normally dist.
```{r}
boxplot(log(AGA$SHLA.G.levels))

```

#### Normality Test
```{r}
shapiro.test(AGA$SHLA.G.levels)
```

#### Distribution plot
```{r}
ggplot(AGA, aes(x = AGA$SHLA.G.levels)) +geom_density()
```

#### Using log transformation
```{r}
ggplot(AGA, aes(x = log(AGA$SHLA.G.levels))) + geom_density()

```

#### Bar plot using ggplot2 library
```{r}
p <- ggplot(AGA, aes(x = POG, y = SHLA.G.levels)) + 
  geom_bar(stat = "identity") +
  labs(title = "Bar plot Of POG vs Levels")
p

```

#### Q-Q plot (or quantile-quantile plot) draws the correlation between a given sample and the normal distribution. 
```{r}
ggqqplot(AGA$SHLA.G.levels)

```

#### QQplot after log transformation
```{r}
ggqqplot(log(AGA$SHLA.G.levels))

```

#### Scatter Plot AGA VS HLA.G.levels
```{r}
plot(AGA$POG,AGA$SHLA.G.levels)

```

#### Ploting Centile Curves using gamlass model
```{r}
h <- gamlss(log(SHLA.G.levels)~cs(POG), sigma.formula = ~cs(POG), 
            nu.formula = ~cs(POG),
            family = BCT, data = AGA)

```


```{r}
centiles(h,xvar = log(AGA$POG), cent = c(50),col.cent=c(4), 
         lwd.cent=c(5) , xlab = "POG", ylab = "HAG Levels", main = "Centile curve For AGA")

```

### Loading SGA Data
```{r}
SGA <- read.csv("C:/Users/DELL/Downloads/SGA1.csv")
head(SGA)

```

```{r}
SGA$Group <- "SGA"
head(SGA)

```

```{r}
names(SGA) <- c("Enrid","POG","SHLA.G.levels","Vist","Group")
head(SGA)

```
### EDA Process
#### Missing Values Analysis and Removing
```{r}
sum(is.na(SGA$SHLA.G.levels))
SGA <- na.omit(SGA)

```

#### Removing outliers
```{r}
boxplot(SGA$SHLA.G.levels)
boxplot(SGA$SHLA.G.levels)$out
boxplot(SGA$SHLA.G.levels, plot=FALSE)$out  
outliers <- boxplot(SGA$SHLA.G.levels, plot=FALSE)$out  
SGA[which(SGA$SHLA.G.levels %in% outliers),]
SGA <- SGA[-which(SGA$SHLA.G.levels %in% outliers),]

```

#### Box plot after removing outliers
```{r}
boxplot(SGA$SHLA.G.levels)

```

#### Log transformation as data is not normally dist.
```{r}
boxplot(log(SGA$SHLA.G.levels))

```

#### Normality Test
```{r}
shapiro.test(SGA$SHLA.G.levels)

```

#### Distribution plot
```{r}
ggplot(SGA, aes(x = SGA$SHLA.G.levels)) +geom_density()

```

#### Using log transformation
```{r}
ggplot(SGA, aes(x = log(SGA$SHLA.G.levels))) + geom_density()
```

#### Bar plot using ggplot2 lib.
```{r}
p <- ggplot(SGA, aes(x = POG, y = SHLA.G.levels)) + 
  geom_bar(stat = "identity") +
  labs(title = "Bar plot Of POG vs HLA.G.Levels")
p

```

#### Q-Q plot (or quantile-quantile plot) draws the correlation between a given sample and the normal distribution. 
```{r}
ggqqplot(SGA$SHLA.G.levels)

```

#### Q-Q plot after log transformation
```{r}
ggqqplot(log(SGA$SHLA.G.levels))
```

#### Scatter Plot
```{r}
plot(SGA$POG,SGA$SHLA.G.levels)

```

#### Ploting Centile Curves using gamlass model
```{r}
h <- gamlss(log(SHLA.G.levels)~cs(POG), sigma.formula = ~cs(POG), 
            nu.formula = ~cs(POG),
            family = BCT, data = SGA)
```

```{r}
centiles(h,xvar = log(SGA$POG), cent = c(50),col.cent=c(4), 
         lwd.cent=c(5) , xlab = "POG", ylab = "HAG Levels", main = "Centile curve For SGA")

```

#### Combining AGA and SGA and use this data for model building
```{r}
pro <- rbind(AGA,SGA)
head(pro)

```

```{r}
tail(pro)
```

```{r}
dim(pro)
```

### Mixed Model

#### There is need to scale POG before building the model. 
```{r}
days <- scale(pro$POG)
pro$days1 <- days[,1]

```

#### Building model1 without using Group 
```{r}
model1 <- lmer(log(SHLA.G.levels)~ days1 + I(days1^2) + 
                I(days1^3)+ (1|Enrid), data = pro, REML = F)
summary(model1)

```

#### ANOVA
```{r}
anova(lmer(log(SHLA.G.levels)~ days1 + I(days1^2) + 
             I(days1^3)+ (1|Enrid), data = pro, REML = F))

```

#### Building model2 with Group as fixed effect
```{r}
model2 <- lmer(log(SHLA.G.levels)~ days1 + I(days1^2) + 
                      I(days1^3) + Group + (1|Enrid), data = pro, REML = F)
summary(model2)

```

#### ANOVA
```{r}
anova(lmer(log(SHLA.G.levels)~ days1 + I(days1^2) + 
             I(days1^3) + Group + (1|Enrid), data = pro, REML = F))

```

#### Comparing model1 and model2 with ANOVA
```{r}
anova(model1,model2)
```

#### Model without POG
```{r}
model3 <- lmer(log(SHLA.G.levels)~ (1|Enrid), data = pro, REML = F)

```

```{r}
summary(model3)
```